Synpotic Soil Characterization

### 2024-07-30 3-reports/manuscript_figures.Rmd
overall drivers. permanova. analyte by region.

@@

ggbiplot::ggbiplot(pca_cluster$pca_int, obs.scale = 1, var.scale = 1,
                   groups = as.character(pca_cluster$grp$cluster), 
                   ellipse = F, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_convexhull(aes(group = pca_cluster$grp$cluster,
                      fill = pca_cluster$grp$cluster), alpha = 0.2)+
  geom_point(size=4,stroke=1, alpha = 1,
             aes(shape = pca_cluster$grp$region,
                 fill = groups), color = "black")+ 
  scale_shape_manual(values = c(23, 21))+
  scale_fill_manual(values = pnw_palette("Sailboat", 3))+
  labs(fill="cluster",
       color = "cluster",
       shape = "region",
       title = "Overall PCA, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

@@

@@

data_combined_wide %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
  ggplot(aes(x = transect, y = Fe))+
  geom_violin(aes(fill = transect), alpha = 0.4)+
  geom_jitter(width = 0.1)+
  geom_text(data = tibble(x = c("upland", "transition", "wc"),
                          label = c("B", "A", "B")),
            aes(x = x, y = 800, label = label), size = 5)+
  labs(x = "",
       y = "Extractable Fe, μg/g")+
  scale_fill_manual(values = pal_transect)+
  theme(legend.position = "none")

data_combined_wide %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
  ggplot(aes(x = transect, y = P))+
  geom_violin(aes(fill = transect), alpha = 0.4)+
  geom_jitter(width = 0.1)+
  geom_text(data = tibble(x = c("upland", "transition", "wc"),
                          label = c("B", "A", "B")),
            aes(x = x, y = 450, label = label), size = 5)+
  labs(x = "",
       y = "Extractable P, μg/g")+
  scale_fill_manual(values = pal_transect)+
  theme(legend.position = "none")

>>>>

overall = 
  pc_all %>% 
  filter(region == "combined") %>% 
  mutate(variable = fct_reorder(variable, abs(PC1))) %>% 
  ggplot(aes(x = abs(PC1), y = variable))+
  geom_point(size = 5, stroke = 1, aes(shape = region, fill = region))+
  scale_fill_manual(values = c("#16879C", "#BB281E" ))+
  scale_shape_manual(values = c(23, 21))+

@@

ggbiplot::ggbiplot(pca_cluster$pca_int, obs.scale = 1, var.scale = 1,
                   groups = as.character(pca_cluster$grp$cluster), 
                   ellipse = F, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_convexhull(aes(group = pca_cluster$grp$cluster,
                      fill = pca_cluster$grp$cluster), alpha = 0.2)+
  geom_point(size=4,stroke=1, alpha = 1,
             aes(shape = pca_cluster$grp$region,
                 fill = groups))+ 
  scale_shape_manual(values = c(23, 21))+
  scale_color_manual(values = pnw_palette("Sailboat", 3))+
  scale_fill_manual(values = pnw_palette("Sailboat", 3))+
  labs(fill="cluster",
       color = "cluster",
       shape = "region",
       title = "Overall PCA, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

@@

## 1b. PERMANOVA

```{r}
library(vegan)

adonis2((data_wide_PCA %>% dplyr::select(where(is.numeric)) %>% drop_na()) ~ 
         (region + transect + horizon + site)^2, 
       data = data_wide_PCA %>% drop_na)

@@

data_combined_wide %>% 
  filter(region == "CB") %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
  ggplot(aes(x = transect, y = Fe))+
  geom_violin(aes(fill = transect), alpha = 0.4)+
  geom_jitter(width = 0.1)+
  geom_text(data = tibble(x = c("upland", "transition", "wc"),
                          label = c("B", "A", "B")),
            aes(x = x, y = 800, label = label), size = 5)+
  labs(x = "",
       y = "Extractable Fe, μg/g")+
  facet_wrap(~region)+
  scale_fill_manual(values = pal_transect)+
  theme(legend.position = "none")

data_combined_wide %>% 
    filter(region == "WLE") %>% 
  mutate(transect = factor(transect, levels = c("upland", "transition", "wc"))) %>% 
  ggplot(aes(x = transect, y = P))+
  geom_violin(aes(fill = transect), alpha = 0.4)+
  geom_jitter(width = 0.1)+
  geom_text(data = tibble(x = c("upland", "transition", "wc"),
                          label = c("B", "A", "B")),
            aes(x = x, y = 450, label = label), size = 5)+
  labs(x = "",
       y = "Extractable P, μg/g")+
  facet_wrap(~region)+
  scale_fill_manual(values = pal_transect)+
  theme(legend.position = "none")
